---
title: "Introduction to R"
author: "Ben Marconi"
date: "February 18, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)
```

Introduction to R Studio and Analysis/Plotting in R

What do the different windows do/mean?

Basic R Commands
```{r}
10^3
(2020-2012)/(2020-1994)
```
Variable names must start with a letter. Naming Variables a user call and manipulate them later
```{r}
percent_of_life_as_academic_slave = (2020-2012)/(2020-1994)
x = percent_of_life_as_academic_slave
```
Simple arithmetic in R can be perfomred as follows using +,-,*,/,%%.%/%
```{r}
x*2

x=5
y=10

x-y
x*y
y/x
y%%x #Modulus (remainder from division)
y%/%x #Interger division
```
Relational Operators: </>/<=/>=/!=/== can be used to compare values of variables 
```{r}
x=5
y=16

x<y
x>y
x<=y
y>=17
y==16
x!=5
```
Getting Sendy with Vectors: Creating short vectors can be accomplished by concatenating integers and numerics.
```{r}
x = c(2,6,3)
y = c(6,4,2)

x+y #operations occur element-wise

x <- c(2,1,8,3)
y <- c(9,4)  
x+y #recycling occurs
x-1
x + c(1,2,3) #no way to recycle, longer vecotr is not intergal multiple of shorter vector


rm(list=ls()) #clears active environment of variables
```
Setting a Working Directory: Choose a folder where data is saved and r scripts can be written with setwd(). 
```{r}
setwd("~/Utah PhD/3rd Year/Spring 2020/R Group/intro_activity")
```
Setting Up a Session for Analysis: install.packages() loads packages not already loaded in base-R to increase the language's functionality. Once a package is installed, calling the package with require() or library() loads the package's data for use. 
```{r}
require(stats)
require(graphics)
require(ggplot2)
```
Some data has already been loaded into R for us. Most of these datasets are clean and behave nicely for example analyses. Data() will pull up a page where you can view the datasets available in R. 
```{r}
data() #explore preloaded datasets in R
data("faithful")
faithful
```
Using the head() function is useful, as it gives you a preview of the dataset without showing (unreasonably in this the case) the entire data frame. 
```{r}
head(faithful)
summary(faithful)
```
The $ can be used to select a column of data within a data frame (the naming convention for how R stores data in a table)
```{r}
duration = faithful$eruptions
range(duration)
mean(duration)
```
Let's do some simple organization of the data with the seq() function to break the data into cleaner intervals.
```{r}

breaks = seq(1.5, 5.5, by = 0.5) #organizing the data into sequence by interval of 0.5
breaks

duration.cut = cut(duration, breaks, right = FALSE) #this classifies the eruptions by interval - FALSE sets intervals closed on left, open on right
duration.frequency = table(duration.cut) #calculate freq of eruption in each sub-interval
duration.frequency

cbind(duration.frequency) #creates a column from table

duration = faithful$eruptions

#Make a PLot
hist(duration, right=FALSE)
colors = c("red", "yellow", "green", "violet", "orange")

waiting = faithful$waiting
head(cbind(duration,waiting)) #Pair eruptions and waiting values to match (x,y) coord
plot(duration, waiting,
     xlab = "Eruption Time",
     ylab = "Time Waited",
     main = "Old Faithful Eruptions",
     sub= "Plots Are Cool")
abline(lm(waiting ~ duration))
fit = lm(waiting ~ duration)
summary(fit)

##Descriptive Statistics
#Hmisc
#pastecs
#psych (uses grouping variable as organizational metric)
#doBy

##LEleL uP KiDdOs 
setwd("~/Utah PhD/3rd Year/Spring 2020/R Group/intro_activity/Rgraphics/dataSets")
install.packages("tidyverse")
library(tidyverse)

housing = read.csv("landdata-states.csv")
head(housing[1:5]) #returns columns 1-5

#Histogram
hist(housing$Home.Value) 

library(ggplot2)
ggplot(housing, aes(x = Home.Value)) + 
  geom_histogram()

plot(Home.Value ~ Date,
     col = factor(State),
     data = filter(housing, State %in% c("MA", "TX")))
legend("topleft",
       legend = c("MA", "TX"),
       col = c("black", "red"),
       pch = 1)

ggplot(filter(housing, State %in% c("MA", "TX")),
       aes(x=Date,
           y=Home.Value,
           color=State))+
  geom_point()+
  geom_line()+
  geom_smooth()

#Diamond Example
library(ggplot2)
gg <- ggplot(diamonds, aes(x=carat, y=price)) 
gg + geom_point()

gg + geom_point(size=1, shape=1, color="steelblue", stroke=2)

gg + geom_point(aes(size=carat, shape=cut, color=color, stroke=carat)) 

gg1 <- gg + geom_point(aes(color=color))
gg2 <- gg1 + labs(title="Diamonds", x="Carat", y="Price")  # ggtitle("title") also changes the title.
print(gg2)
```